---------------------------------------------------------------------------
title: "Assignment 1 - Reproducible Research"
author: "Satya"
# This assignment makes use of data from a personal activity monitoring device.

date: "December 17, 2016"
 
------------------------------------------------------------------

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

## R Markdown

#1.Code for reading in the dataset and/or processing the data.

```{r echo=TRUE}

library(lattice)
library(dplyr)
library(knitr)
library(ggplot2)
setwd("C:\\R\\Reproducable\\week2")
dataActivity <- as.data.frame(read.csv(file = "activity.csv", header = TRUE))
dataActivity$Tdate = as.Date( dataActivity$date)
dataActivity$steps = as.numeric( dataActivity$steps)

```


#1. Calculate Average total number of steps taken per day


```{r  echo=TRUE}

stepsPerDay <- dataActivity %>%
               group_by ( Tdate) %>%
               summarise( totalSteps = sum(steps) ) 


```

#2. Histogram of the total number of steps taken each day


```{r  echo=TRUE}

with(stepsPerDay, hist(stepsPerDay$totalSteps, 
                       xlim = c(0, 25000), ylim = c(0, 30),
                       main="Total number of steps take per day", 
                       col="green", 
                       xlab="Steps taken"))

```


#3. Mean and median number of steps taken each day

```{r  echo=TRUE}

meanSteps <- mean(stepsPerDay$totalSteps, na.rm = TRUE)
meanSteps

medianSteps <- median(stepsPerDay$totalSteps, na.rm = TRUE)
medianSteps

```


#4. Time series plot of the average number of steps taken


```{r  echo=TRUE}

stepsPerInterval <- dataActivity %>%
  group_by ( interval) %>%
  summarise( totalSteps = mean(steps, na.rm = TRUE) ) 

 PLot1 <- ggplot(stepsPerInterval, aes(interval, totalSteps)) + geom_line(colour = "steelblue4", 
       lwd = 1) + labs(title = "Average Number of Steps Taken per Interval")


 print(PLot1)
 
```

# 5. The 5-minute interval that, on average, contains the maximum number of steps
 

```{r  echo=TRUE}

 maxStep <- stepsPerInterval$interval[ which.max(stepsPerInterval$totalSteps)]
 
maxStep

max(stepsPerInterval$totalSteps)

```

#6. Code to describe and show a strategy for imputing missing data.

#6.1  Total Number of missing Values.

```{r  echo=TRUE}

 totalMissingValueRows <-  sum(is.na(dataActivity))
 
 totalMissingValueRows 
 

```


#6.2  Filling in the missing values using the mean of the 5-minutes interval


```{r  echo=TRUE}


 for (i in 1:length(dataActivity$steps)) {
   if (is.na(dataActivity[i, 1])) {
   mean1 <- mean( dataActivity$steps, na.rm = TRUE)
  
    dataActivity[i, 1] <- mean1
     
   } else {
    dataActivity[i, 1] <- dataActivity[i, 1]
   }
   dataActivity
 }
 


```

#6.3  New dataset with the missing values filled in

```{r  echo=TRUE}

head(dataActivity)
 
 tail(dataActivity)

```

#7.  Histogram of the total number of steps taken each day after missing values are imputed
 

```{r  echo=TRUE}

 stepsPerDay <- dataActivity %>%
   group_by ( Tdate) %>%
   summarise( totalSteps = sum(steps) ) 
 
 with(stepsPerDay, hist(stepsPerDay$totalSteps, 
                        xlim = c(0, 25000), ylim = c(0, 30),
                        main="Total number of steps take per day", 
                        col="red", 
                        xlab="Steps taken"))
 
 

```

#7.1  Mean and median of the total number of steps taken per day
 


```{r  echo=TRUE}
meanSteps <- mean(stepsPerDay$totalSteps, na.rm = TRUE)
 meanSteps
 
 medianSteps <- median(stepsPerDay$totalSteps, na.rm = TRUE)
 medianSteps

```

#7.2  The variation of the mean and the median between before and after imputing the missing values is small for this sample.
 
 
 #8.  Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
 
 #8.1  Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.


```{r  echo=TRUE}

 dataActivity$date <- strptime(dataActivity$date, "%Y-%m-%d")
 
 dataActivity$date <- as.POSIXct(dataActivity$date)
 
 dataActivity <- mutate(dataActivity, day = weekdays(date))

  dataActivity[dataActivity$day %in% c("Friday", "Monday", "Thursday", "Tuesday" ,"Wednesday"), "day"] <- "weekdays"
 
  dataActivity[dataActivity$day %in% c( "Saturday", "Sunday") , "day"] <- "weekends"
 
  dataActivity$day <- as.factor(dataActivity$day)

  head(dataActivity) 
```

#8.2  Panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days.


```{r  echo=TRUE}
stepsPerInterval <- dataActivity %>%
  group_by ( day, interval) %>%
  summarise( totalSteps = mean(steps, na.rm = TRUE) ) 


ggplot(stepsPerInterval, aes(interval, totalSteps)) + geom_line(color = "steelblue4", lwd = 2) + 
  facet_wrap(~day, ncol = 1) + labs(title = "Average Number of Steps Taken")

 
```
 